---
title: "PML_project"
author: "Sergio Rodriguez"
date: "23 de junio de 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

## Data Loading
```{r}
suppressWarnings(library(dplyr))
suppressWarnings(library(ggplot2))
suppressWarnings(library(lubridate))
suppressWarnings(library(caret))
suppressWarnings(library(randomForest))
suppressWarnings(library(rpart))
suppressWarnings(library(rpart.plot))
suppressWarnings(library(corrplot))

data.train<- read.csv("C:/Users/Sergio Rodriguez/Documents/Coursera/Practical Machine Learning/pml-training.csv", na.strings = c("NA", "#DIV/0!", ""))

data.test<- read.csv("C:/Users/Sergio Rodriguez/Documents/Coursera/Practical Machine Learning/pml-testing.csv", na.strings = c("NA", "#DIV/0!", ""))
```

## Data Transformation

Understand the data, convert date and add a new varaible day.
```{r}
dim(data.train)
data.train$cvtd_timestamp<- as.Date(data.train$cvtd_timestamp, format = "%m/%d/%Y %H:%M")
data.train$Day<-factor(weekdays(data.train$cvtd_timestamp)) #Add day variable

table(data.train$classe)
prop.table(table(data.train$classe)) 
prop.table(table(data.train$user_name))
prop.table(table(data.train$user_name,data.train$classe),1) 
prop.table(table(data.train$user_name,data.train$classe),2) 
prop.table(table(data.train$classe, data.train$Day),1) 
qplot(x=Day, fill=classe, data = data.train)
```

## Data Cleaning
```{r}
# Remove columns with NA missing values
data.train <- data.train[, colSums(is.na(data.train)) == 0]
data.test <- data.test[, colSums(is.na(data.test)) == 0] 

# Remove columns that are not relevant to accelerometer measurements.
classe<- data.train$classe
trainRemove<- grepl("^X|timestamp|window", names(data.train))
data.train<- data.train[, !trainRemove]
trainCleaned<- data.train[, sapply(data.train, is.numeric)]
trainCleaned$classe<- classe
testRemove<- grepl("^X|timestamp|window", names(data.test))
data.test<- data.test[, !testRemove]
testCleaned<- data.test[, sapply(data.test, is.numeric)]
```

## Create Train and Test Data Sets
```{r}
set.seed(22519)
inTrain <- createDataPartition(trainCleaned$classe, p=0.70, list=F)
trainData <- trainCleaned[inTrain, ]
testData <- trainCleaned[-inTrain, ]
```

## Data Modelling
```{r}
controlRf <- trainControl(method="cv", 5)
rfmod<- train(classe ~., data=trainData, method="rf", trControl=controlRf, importance=TRUE, ntree=100)
rfmod
```

## Accuracy of Model
```{r}
predictRfmod<- predict(rfmod, testData)
confusionMatrix(testData$classe, predictRfmod)

accuracy <- postResample(predictRfmod, testData$classe)
accuracy

Error <- 1 - as.numeric(confusionMatrix(testData$classe, predictRfmod)$overall[1])
Error
```
So, the estimated accuracy of the model is 99.32% and the estimated out-of-sample error is 0.68%.

## Predicting on test Data
```{r}
result <- predict(rfmod, testCleaned[, -length(names(testCleaned))])
result
```

### Correlation MAtrix
```{r}
corrPlot <- cor(trainData[, -length(names(trainData))])
corrplot(corrPlot, method="circle")
```

### Tree Visualization
```{r}
rtree<- rpart(classe ~ ., data=trainData, method="class")
prp(rtree)
```